Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2018-02-12 aggregation/agg-community.tex
Tiny tweaks. I should really put this down for awhile.

To advance LCA practice, the community should embrace the notion of a software-independent PSM description and consider how to best achieve it.  The SETAC North America roadmap module for PSM description and revision \citep{Kuczenski_JLCA_2018} includes milestones toward this goal.  The disclosure framework presented here is offered as a candidate for Milestone 2.1, ``a minimal description of a PSM.''  The community should give this proposal a robust critical evaluation, and researchers should consider how to conceptualize their own models in terms of the framework.  Other milestones and cross-cutting issues are of crucial interest:
\begin{enumerate}[label=(\alph*)]
\item LCI database providers must provide their users with unambiguous, static, resolvable references to background datasets (Milestone 1.1) so that their use in a model may be documented precisely (Milestone 3.1);
\item The transformation of multi-functional processes to single-function processes should be described using the framework (Milestones 1.2 and 1.4);
\item Software makers should enable their users to automatically generate lists of foreground, background, and emission flows a model foreground (disclosure items \ref{itm:fg} through \ref{itm:em}) and to export foreground matrices (disclosure items \ref{itm:af} through \ref{itm:bf}) from a model in a concise format (Milestones 2.2 and 3.4);
\item During ISO critical review, practitioners and reviewers should strive to come to agreement on the formal structure and contents of the model being reviewed;
\item The LCA research community, including interested parties within the International Society for Industrial Ecology, the UNEP/SETAC Life Cycle Initiative, and other organizations, should collectively pursue consistent, shared definitions of quantities of measure, flows, and contexts used in LCIA.  %These definitions should be published and curated using Semantic Web techniques and practices.
LCIA method developers should immediately begin using these consensus definitions instead of their own internal lists when publishing flow characterization data, so that software makers can interpret the publications correctly.  %A characterization of a flow defined privately, into a context defined privately, cannot be interpreted meaningfully by the public.
\end{enumerate}
As these milestones are pursued and achieved, it will be possible for a standalone description of a PSM to be computed and extended by another party, regardless of the software used to construct the model originally.  This will tremendously improve the capacity for critical review and reuse of LCA models.

>>>>

To advance LCA practice, the community should embrace the notion of a software-independent PSM description and consider how to best achieve it.  The SETAC North America roadmap module for PSM description and revision \citep{Kuczenski_JLCA_2018} includes milestones toward this goal.  The disclosure framework presented here is offered as a candidate for Milestone 2.1, ``a minimal description of a PSM.''  The community should give this proposal a robust critical evaluation, and researchers should consider how to conceptualize their own models in terms of the framework.  Other milestones and cross-cutting issues are of crucial interest:
\begin{enumerate}[label=(\alph*)]
\item LCI database providers must provide their users with unambiguous, static, resolvable references to background datasets (Milestone 1.1) so that their use in a model may be documented precisely (Milestone 3.1);
\item The transformation of multi-functional processes to single-function processes should be described using the framework (Milestones 1.2 and 1.4; see supporting information);
\item Software makers should enable their users to automatically generate lists of foreground, background, and emission flows a model foreground (disclosure items \ref{itm:fg} through \ref{itm:em}) and to export foreground matrices (disclosure items \ref{itm:af} through \ref{itm:bf}) from a model in a concise format (Milestones 2.2 and 3.4);
\item During ISO critical review, practitioners and reviewers should strive to come to agreement on the formal structure and contents of the model being reviewed;
\item The LCA research community, including interested parties within the International Society for Industrial Ecology, the UNEP/SETAC Life Cycle Initiative, and other organizations, should collectively pursue consistent, shared definitions of quantities of measure, flows, and contexts used in LCIA.  %These definitions should be published and curated using Semantic Web techniques and practices.
LCIA method developers should immediately begin using these consensus definitions instead of their own internal lists when publishing flow characterization data, so that software makers can interpret the publications correctly.  %A characterization of a flow defined privately, into a context defined privately, cannot be interpreted meaningfully by the public.
\end{enumerate}
As these milestones are pursued and achieved, it will be possible for a standalone description of a PSM to be computed and extended by another party, regardless of the software used to construct the model originally.  This will tremendously improve the capacity for critical review and reuse of LCA models.

### 2018-02-12 aggregation/agg-pagg.tex
Tiny tweaks. I should really put this down for awhile.

The $A_d$ and $B_f$ matrices are then partitioned into two components that sum to the original:
\begin{equation}\begin{aligned}
  A_d =\; & A_{d,pub} + A_{d,priv} \\
  B_f =\; & B_{f,pub} + B_{f,priv}
  \label{eqn:partition}
  \end{aligned}
  \end{equation}
This is illustrated in Fig.~\ref{fig:private}b.  Substituted back into Eq.~\ref{eqn:study}, the private portions are aggregated into a private life cycle inventory vector $\mathbf{b}_{priv}$, while the public portions remain disaggregated:
\begin{align}
  \mathbf{b}_{priv} =\; & (B_{f,priv} + B_x\times A_{d,priv})\times \tilde{\mathbf{x}}\\
  s =\; & \mathbf{c}^T\times\left((B_{f,pub} + B_{f,priv}) + B_x\times (A_{d,pub} + A_{d,priv})\right)\times\tilde{\mathbf{x}}\\
  =\; & \mathbf{c}^T\times\left((B_{f,pub} + B_x\times A_{d,pub})\times\tilde{\mathbf{x}} + \mathbf{b}_{priv}\right)\label{eqn:private}
\end{align}
Any foreground nodes that are completely contained within the private partition can be omitted from the disclosed foreground.  The aggregation result can then be included as a separate foreground node with a unit weight, and the aggregated inventory vector included as the corresponding column of $B_f$.  This is illustrated in Fig.~\ref{fig:private}c.

>>>>

The $A_d$ and $B_f$ matrices are then partitioned into two components that sum to the original:
\begin{equation}\begin{aligned}
  A_d =\; & A_{d,pub} + A_{d,priv} \\
  B_f =\; & B_{f,pub} + B_{f,priv}
  \label{eqn:partition}
  \end{aligned}
  \end{equation}
This is illustrated in Fig.~\ref{fig:private}b.  Substituted back into Eq.~\ref{eqn:study}, the private portions are aggregated into a private life cycle inventory vector $\mathbf{b}_{priv}$:
\begin{equation}
  \mathbf{b}_{priv} = (B_{f,priv} + B_x\times A_{d,priv})\times \tilde{\mathbf{x}}
\end{equation}
while the public portions remain disaggregated:
\begin{align}
  s =\; & \mathbf{c}^T\times\left((B_{f,pub} + B_{f,priv}) + B_x\times (A_{d,pub} + A_{d,priv})\right)\times\tilde{\mathbf{x}}\\
  =\; & \mathbf{c}^T\times\left((B_{f,pub} + B_x\times A_{d,pub})\times\tilde{\mathbf{x}} + \mathbf{b}_{priv}\right)\label{eqn:private}
\end{align}
Any foreground nodes that are completely contained within the private partition can be omitted from the disclosed foreground.  The aggregation result can then be included as a separate foreground node with a unit weight, and the aggregated inventory vector included as the corresponding column of $B_f$.  This is illustrated in Fig.~\ref{fig:private}c.

### 2018-02-12 aggregation/agg-semantic.tex
Tiny tweaks. I should really put this down for awhile.

In practice, while it is easy to reproduce a set of sparse matrices, there is considerable potential for ambiguity in identifying the background datasets and emissions to which those matrix entries correspond.
In order to achieve the transparency and reproducibility promised by this disclosure framework, it must be possible for study authors and readers to agree on the meaning of the references contained in items~\ref{itm:bg} and \ref{itm:em}.
The FAIR guidelines, which were designed to help scientists organize data for Findability, Accessibility, Interoperability, and Reusability \citep{Wilkinson_2016}, provide valuable perspective on this goal.

Central to many of the FAIR guiding principles is the use of linked semantic data \citep{Bizer_2009}, in which an object to be interpreted is signified by a link to a resource on the World Wide Web, typically referred to as a Uniform Resource Identifier (URI) or a hyperlink.  This link serves as a unique identifier and a resolvable reference to metadata, and the linked content can be  curated by the study authors and other members of the community \citep{Khan_2011}.

>>>>

In practice, while it is easy to reproduce a set of sparse matrices, there is considerable potential for ambiguity in identifying the background datasets and emissions to which those matrix entries correspond.
In order to achieve the transparency and reproducibility promised by this disclosure framework, it must be possible for study authors and readers to agree on the meaning of the references contained in items~\ref{itm:bg} and \ref{itm:em}.
The FAIR guiding principles, which were designed to help scientists organize data for Findability, Accessibility, Interoperability, and Reusability \citep{Wilkinson_2016}, provide valuable perspective on this goal.

Central to many of the FAIR guidelines is the use of linked semantic data \citep{Bizer_2009}, in which an object to be interpreted is signified by a link to a resource on the World Wide Web, typically referred to as a Uniform Resource Identifier (URI) or a hyperlink.  This link serves as a unique identifier and a resolvable reference to metadata, and the linked content can be  curated by the study authors and other members of the community \citep{Khan_2011}.
